---
title: "Three type birth-mutation-death process"
author: "Itziar Irurzun Arana"
date: "`r Sys.Date()`"
output:  
  prettydoc::html_pretty:
    theme: cayman 
    highlight: github
vignette: >
  %\VignetteIndexEntry{Three type birth-mutation-death process}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warnings= FALSE
)
```
Load library:
```{r}
library(ACESO)
```

## Data
In the current evaluation cell viability data resulting from the exposure of BT-20 triple-negative breast cancer cell line  to different concentration values of two small molecule kinase inhibitors (Dactolisib and Trametinib) is analyzed.
```{r}
data(Dactolisib_Trametinib_combination)
head(Dactolisib_Trametinib_combination)

DrugA=as.character(Dactolisib_Trametinib_combination$Small.Molecule.Name[1])
DrugB=as.character(Dactolisib_Trametinib_combination$Small.Molecule.2.Name[1])
print(c(DrugA,DrugB))
```

Exploratory analysis of the data:

```{r}
library(ggplot2)
ggplot(data=Dactolisib_Trametinib_combination,aes(x=(CONC),y=Viable.cells,col=factor(CONC2)))+
  geom_point(size=1.5)+
  xlab(paste0("[",DrugA,"] (µM)"))+scale_colour_discrete(name=paste0("[",DrugB,"] (µM)"))+
  theme_classic()+
  theme(text = element_text(size=14))
```
BT-20 cells sensitive to both drugs were defined as type 0 cells, whereas cells resistant to trametinib and dactolisib were defined as type 1 and type 2 cells respectively. 

# Dynamics of sensitive cells:

##Calculate net growth rate
Calculate the net growth rates of the BT-20 cell type for every concentration of the two drugs. Use the **net_growth_rate** function for this purpose, where an exponential growth function is assumed to calculate those rates.
```{r}
growth_data<-net_growth_rate(Dactolisib_Trametinib_combination)
head(growth_data)
```
Plot the results:
```{r}
ggplot(data=growth_data,aes(x=(CONC),y=Net_growth,col=factor(CONC2)))+geom_point()+geom_line(linetype="dashed")+
  theme(text = element_text(size=14))+xlab(paste0("[",DrugA,"] (µM)"))+
  scale_colour_discrete(name=paste0("[",DrugB,"] (µM)"))+
  theme_classic()+ylab("Net growth (1/h)")+theme(text = element_text(size=14))
```
## Calculate birth and death rates
As in the LINCS database there is no information regarding apoptosis assays available to estimate the death rate of type 0 cells (d0) and assuming that these targeted therapies do not induce cell death per se, a different d0 value was defined for each combination of the drugs, which was equal to the minimum value needed to ensure that all the birth rates of type 0 cells (b0) were positive regardless of the treatment conditions:
```{r}
d0=vector("list", length = length(unique(growth_data$Cell.line)))
names(d0)=unique(as.character(growth_data$Cell.line))
d0[[1]]=-min(growth_data$Net_growth)
d0
```
We use the **net_growth_rate** function again, to not only calculate the net growth rate parameter of the cells but also the birth and death rates:
```{r}
growth_data<-net_growth_rate(growth_data,death_rate = d0)
ggplot(data=growth_data,aes(x=(CONC),y=Birth_rate,col=factor(CONC2)))+geom_point()+geom_line(linetype="dashed")+
  theme(text = element_text(size=14))+xlab(paste0("[",DrugA,"] (µM)"))+
  scale_colour_discrete(paste0("[",DrugB,"] (µM)"))+
  theme_classic()+ylab("Birth rate of \n sensitive cells, b0 (1/h)")+theme(text = element_text(size=14))

```
Another plot type to analyze the response surface of the data:
```{r}
#We are going to remove the columns that are not necessary for the analysis of the data in this example:
GD=growth_data[,c('Cell.line','CONC','CONC2','Net_growth','Type','Birth_rate','Death_rate')]
GD=unique(GD)
```
Use **responseMap** function to reorganize the data to easily plot the surface of the birth rate values versus for both drug concentrations:
```{r}
rmap <- responseMap(Birth_rate~CONC+CONC2,GD,logscale=T,interpolate=FALSE)
DifferenceSurface.plot(rmap,zcenter=max(growth_data$Birth_rate)/2,
                       xl=paste0("[",DrugA,"] (µM)"),
                       yl=paste0("[",DrugB,"] (µM)"),
                       zl="Birth rate of \n sensitive cells, b0 (1/h)",
                       mid="yellow",low="hotpink1",high="darkturquoise")

```
Use **responseMap** again and **plot.ResponseSurface** to make a contour plot of the data. For this plot a interpolation of the data is needed (see logscale=T in the **responseMap** function):
```{r}
rmap <- responseMap(Birth_rate~CONC+CONC2,GD,logscale=T)

ResponseSurface.plot(rmap,xl=paste0("[",DrugA,"] (µM)"),
                     yl=paste0("[",DrugB,"] (µM)"),
                     zl="Birth rate of \n sensitive cells, b0 (1/h)",
                     palette=c("hotpink1","yellow","darkturquoise"))

```
## Sensitive cell model fitting
Birth rate parameters were fitted nonparametrically using a Generalized Additive Model (GAM) to generate the predicted surface required for the simulation of the evolutionary process:
```{r}
gam.model=Multiple.resp.surface.fit(GD, title=", GAM",Drug1.name=paste0("[",DrugA,"] (µM)"),Drug2.name=paste0("[",DrugB,"] (µM)"))
```

# Dynamics of resistant cells: