---
title: "One drug resistance mechanisms"
author: "Itziar Irurzun Arana"
date: "`r Sys.Date()`"
output:  
  prettydoc::html_pretty:
    theme: cayman 
    highlight: github
vignette: >
  %\VignetteIndexEntry{Complex dosing schedules}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warnings= FALSE
)
```
Load library:
```{r}
library(ACESO)
```
### Read cell proliferation file:

Use **read.cellcount.data** function to create a data.frame with the info about cell proliferation under different drug concentrations.

Mandatory columns to be in the .csv file: 

* Viable.cells: cell count data
* Time: time point
* CONC: drug concentration

Optional columns: 

* Replicate: Technical replicate. If missing the program assumes that there is only 1 replicate.
* Cell.line: Name of the cell line. If missing the fucntion introduces the name 'Cell line 1'
* Type: numerical column specifiying the cell type (0: sensitive cells, 1: resistant cell to drug A, etc.). If missing the program assumes Type=0.

```{r}
growth_data=read.cellcount.data("cell_viability_assay.csv", sep=";")
head(growth_data)
```
